% !TEX root = ../../mat701_notes.tex
\newpage
\section{Repeated Integration}
\subsection{Fubini's Theorem}


Given the function $f(x,y)= \dfrac{1}{\sqrt{xy}}$ on the interval $[0,1] \times [0,1]$, we already have a concept of $\ds\int_{[0,1] \times [0,1]} f$. But how does this integral relate to the integral
	\[
	\int_0^1 \left( \int_0^1 \dfrac{1}{\sqrt{xy}} \; dy \right) \; dx?
	\]
One can ask the same questions and confront the same issues with infinite double summations, e.g. $\ds\sum_{i,j}^\infty a_{ij}$.


\begin{ex} \label{ex:matrixex}
Consider the infinite matrix with 1s on the main diagonal and $-1$ on the superdiagonal and zeros elsewhere. The sum of each row is 0 and, except for the first column,  the sum of column is 0. This implies
	\[
	\begin{split}
	\sum_{i=1}^\infty \left( \sum_{j=1}^\infty a_{ij} \right)&= 0 \\
	\sum_{j=1}^\infty \left( \sum_{i=1}^\infty a_{ij} \right)&= 1
	\end{split}
	\]
One can create a parallel example with integrals. For example, take $f=1,-1,0$ and consider
	\[
	\begin{split}
	\int_0^\infty \left( \int_0^\infty f(x,y) \; dx \right) \; dy&= 0 \\
	\int_0^\infty \left( \int_0^\infty f(x,y) \; dy \right) \; dx&= 0
	\end{split}
	\]
Note that $f \notin L^1([0,\infty) \times[0,\infty))$. 
\end{ex}


The goal of this section is to establish Fubini's Theorem, stated below, and its applications. This was proven by Guido Fubini in 1907. The theorem allows determines conditions under which it is possible to compute double integrals using iterated integrals. Consequently, this allows one to exchange the order of integration. In 1909, a student of Fubini named Leonida Tonelli established a similar theorem (Tonelli's Theorem) but applying to functions which are nonnegative instead of integrable. 


\begin{restatable}[Fubini's Theorem]{thm}{fubini} \label{thm:fubinithm}
Let $I_1 \subseteq \R^n$, $I_2 \subseteq \R^m$ be intervals, and $I:= I_1 \times I_2$. Suppose that $f \in L^1(I)$. Then
	\begin{enumerate}[(i)]
	\item for almost every $\mathbf{x} \in I_1$, the `slice' function $\mathbf{y} \mapsto f(\mathbf{x},\mathbf{y})$ is measurable and integrable on $I_2$.
	\item the `once-integrated' function $\mathbf{x} \mapsto \int_{I_2} f(\mathbf{x},\mathbf{y}) \; d\mathbf{y}$ is measurable and integrable on $I_1$, and its value is $\int_I f$, i.e.
		\[
		\iint_I f(\mathbf{x},\mathbf{y}) \; d\mathbf{x} d\mathbf{y} = \int_{I_1} \left( \int_{I_2} f(\mathbf{x},\mathbf{y}) \; d\mathbf{y} \right) \; d\mathbf{x}.
		\]
	\end{enumerate}
\end{restatable}


The statement `for almost every $\mathbf{x}$' is required as a slice may have `issues' on some set of measure 0, which will not affect the integral. As a parallel issue, for example, consider the matrix in Example~\ref{ex:matrixex}. One could instead insert for some zero entry a nonmeasurable function or the value $\infty$. So while the `slice' function may be problematic, its integral is unchanged since its values were changed on a set of measure zero. Recall that 
	\[
	\int_I f= \int_{\R^{n \times m}} f \cdot \chi_I.
	\]
By replacing $f$ with $f \chi_I$, i.e. taking $f \equiv 0$ outside of $I$, we may reduce the proof of Fubini's Theorem to the case where  $I_1= \R^n$, $I_2= \R^m$, and $I= \R^{n+m}$. Let $\mathcal{F}:= \{ f \in L^1( \R^{n+m}) \colon \text{Fubini holds for } f \}$. To prove Fubini's Theorem, we prove that $\mathcal{F}=L^1$. First, observe the following properties of $\mathcal{F}$:













% Finish below prop


\begin{prop}
Let $\mathcal{F}:= \{ f \in L^1( \R^{n+m}) \colon \text{Fubini holds for } f \}$. Then
        \begin{enumerate}[(i)]
        \item $\mathcal{F}$ is a vector space: $f,g \in \mathcal{F}$, then $\alpha f + \beta g \in \mathcal{F}$.
        \item $\mathcal{F}$ is closed under monotone limits: if $f_k \in \mathcal{F}$ and $f_k \nearrow f$ or $f_k \searrow f$, where $f \in L^1(\R^{n \times m})$, then $f \in \mathcal{F}$.
        \item Squeeze Property: if $g \leq f \leq h$ and $g=h$ a.e., where $g,h \in F$, then $f \in \mathcal{F}$
        \item $\chi_I \in F$ for any closed bounded interval $I$. 
        \end{enumerate}
\end{prop}


\pf \hfill
\begin{enumerate}[(i)]
\item This is immediate from the fact that if $f,g$ are measurable, then $f+g$ and $cf$ are measurable. Similarly, if $f,g \in L(E)$, then $f+g, cf \in L(E)$.

\item By replacing $f_k$ with $-f_k$ for all $k$, we may assume without loss of generality that $f_k \nearrow f$. For each $k$, there is a set $Z_k \subseteq \R^n$ with measure zero such that the $\mathbf{x}$-slice function of $f_k$ is integrable for $\mathbf{x} \notin Z_k$. Let $Z= \cup_k Z_k$. Since $Z$ is a countable union of sets with measure zero, $Z$ has measure zero. If $\mathbf{x} \notin Z$, then the $\mathbf{x}$-slice function $f_k(\mathbf{x},\mathbf{y})$ is integrable for all $k$, and by the MCT applied to $\{f_k(\mathbf{x},\mathbf{y})\}$ as functions of $\mathbf{y}$, we have
	\[
	h_k(\mathbf{x}):= \int f_k(\mathbf{x},\mathbf{y}) \; d\mathbf{y} \nearrow \int f(\mathbf{x},\mathbf{y}) \; d\mathbf{y}=: h(\mathbf{x}).
	\]
By assumption, $h_k(\mathbf{x})$ is integrable as a function of $\mathbf{x}$, $f_k \in L(I)$, and $\ds\iint f_k(\mathbf{x},\mathbf{y}) \; d\mathbf{x} d\mathbf{y}= \int h_k(\mathbf{x}) \; d\mathbf{x}$. But then by the MCT,
	\[
	\iint f(\mathbf{x},\mathbf{y}) \; d\mathbf{x} d\mathbf{y} = \int h(\mathbf{x}) \; dx.
	\]
As $f \in L(I)$, we have $h \in L(I_1)$, implying that $h$ is finite almost everywhere. Therefore, Fubini's Theorem holds for $f$ also. 
\end{enumerate}







We now prove the second. For almost all $x$, the slice of $g$ and $h$ are both integrable, and since the slice of $f$ is between, it is also integrable. 


(Proof of 3???) Once integrated functions for $g, f, h$ satisfy $\int g \; dy \leq \int f \; dy \leq \int h \; dy$ for almost all $x$. Integrate with respect to $x$
	\[
	\left( \int g \; dy \right),
	\]
but then 
	\[
	\left( \int g \; dy \right) \; dx \leq \left( \int f \; dy \right) \; dx  \leq \left( \int h \; dy \right) \; dx 
	\]
But the left/right integrals are equal to $\int_I g$ and $\int_I h$, respectively. But these are equal since $g=h$ a.e.. Therefore, $\left( \int f \; dy \right) \; dx= \int_I f$.


Proof of 4: Slice function of $\chi_I$ is either $\equiv 0$ or $\chi_{I_2}$, so $\int \chi_I= |I_2| \chi_{I_1}$. Then $\int \chi_I \;dy \; dx= |I_2| |I_1|= |I|= \int \chi_I$. \qed \\






















% Fubini: f \in L^1(I_1 \times I_2)
% Tonelli: f \geq 0, measurable on I_1 \times I_2


We are now in a position to prove Fubini's Theorem.

\fubini*

\pf We establish $\mathcal{F}= L^1$. We show this for $G_\delta$ sets and sets of measure 0, which will be enough. 

a. First, we show that $X_I \in \mathcal{F}$ for any closed bounded interval. This does not necessary worked for closed. Indeed, open sets are union of closed sets. 

% Picture: dotted square, the solid squares inside getting closer and closer to full square.

Consider the sequence of characteristic functions on the boxes in the figure, say $\chi_{I_k} \nearrow \chi_I$. Now for a set which is not open or closed, squeeze this between an open set containing and a closed set inside it.

b. Now $\chi_G \in \mathcal{F}$ if $G$ is open and $|G|<\infty$. Tile $G$ by half-open cubes $Q_j$, $\chi_G= \sum \chi_{Q_j}$. Monotone limit of partials sums of 1 and 2 bullets above. 

c. $\chi_{H} \in \mathcal{F}$ if $H$ is a $G_\delta$ set with $|H|<\infty$. Indeed, there exists $G$ open, $H \subset G$, and $|G|<|H|+\infty$ so if $H= \cap_{k=1}^\infty G_k$, $G_k$, with $H= (G \cap G_1) \cap (G\cap G_1 \cap G_2) \cap (G \cap G_1 \cap G_2 \cap G_3) \cap \cdots$ so it is a nested intersection of open sets of finite measure. Then use second bullet above. 

d. Now $\chi_Z \in \mathcal{F}$ if $|Z|=0$. There exists a $G_\delta$ set $H$ such that $Z \subset H$ and $|H|=0$. Now $0 \leq \chi_Z \leq \chi_H$ and $\chi_H=0$ a.e., then by squeeze (bullet 3) have the conclusion. 

e. We now show $\chi_X \in \mathcal{F}$ if $|E|<\infty$, any set of finite measure. Write $E= H \sm Z$. $H$ is $G_\delta$, $|Z|=0$. Then $\chi_E= \chi_H \sm \chi_Z$ and use linearity. 

f. integrable simple functions $\in \mathcal{F}$. Then nonnegative integrable in $\mathcal{F}$. Then integrable in $\mathcal{F}$ since $f= f^+ - f^-$. \qed \\














% ------------------------
%  Tonelli's Theorem
% ------------------------

% Add from textbook.


% Tonelli


To use Fubini, we need $f \in L^1$. A lot of times, this is what we are trying to show! Fubini cannot be used to prove $f \in L^1(I_1 \times I_2)$. Take $f(x,y)= \dfrac{1}{\sqrt{xy}}$ on $I:=[0,1] \times [0,1]$. Cannot say $\int_I f = \int_0^1( \int_0^1 \dfrac{1}{\sqrt{xy}} \; dx)\;dy$ but need $f$ integrable, which is what you are trying to show. For these purposes, we use Tunelli's Theorem.

\begin{thm}
Same as Fubini Theorem, $f \geq 0$, measurable. Suppose $f \geq 0$, measurable on $I_1 \times I_2$. Then 

a): FOr a.e. $x \in I_1$, the slice function $y \mapsto f(x,y)$ is measurable on $I_2$.
b): the once-integrated function $x \mapsto \int_{I_2} f(x,y) \; dy$ is measurable on $I_1$ andits integral is $\int_I f$ (could be $\infty$). 
\end{thm}


In the example above, $\int_0^1 \dfrac{1}{\sqrt{xy}} \; dy= \dfrac{2}{\sqrt{x}}$ is measurable and $\int_0^1 \dfrac{2}{\sqrt{x}} \; dx = 4$ so $\int_I \dfrac{1}{\sqrt{xy}}= 4$. 

\pf Let $f_k= \min(f,k) \chi_{[-k,k]^n}$ integrable and $f_k \nearrow f$. Then Fubini applies. Spelled out. $\int_{I_1 \times I_2} f_k = \int_{I_1} \left( \int_{I_2} f_k(x,y) \; dy \right) \; dx$. The inner integral converegs a.e. y to $\int_{I_2} f(x,y) \; dy$ by MCT. THen left MCT $\int_I f$. And then MCT for outer integral. done. \qed \\





























% ------------------------
%  Applications
% ------------------------
\subsection{Applications of Fubini's Theorem}

Let $fL E \to [0,\infty]$ be measurable. $\omega_f(y)= |\{x \in E \colon f(x)>y\}|$. Then $\int_E f = \int_0^\infty \omega_f(y) \; dy$. 

\pf Recall $R(f,E)= \{(x,y) \in \R^{n+1} \colon x \in E, 0 \leq y \leq f(x)\}$. $\Gamma(f,E)= \{(x,y) \in \R^{n+1} \colon y=f(x)\}$. Now $|\Gamma|-0$, $g= \chi_{R \sm \Gamma}$ so $g(x,y)=1$ if $0 \leq y<f(x)$, 0 otherwise. $\int(\int g(x,y) \; dy)\;dx= \int(\int g(x,y) \; dx) \; dy)$. and $\int f(x) \chi_E \; dx = \int \omega_f(y) \; dy$ so $\int_E f= \int_0^\infty \omega_f(y) \; dy$. This last integral is even Riemann integrable since it is monotone. 



Generalizations: $f$ measurable, $\int_E |f|^p= p \int_0^\infty y^{p-1} \omega_{|f|}(y) \; dy$. $p>0$.

We know $\int_E |f|^p= \int_0^\infty \omega_{|f|^p}(y) \; dy$. Let $y=z^p$. Then this is $p \int_0^\infty z^{p-1} \omega_{|f|^p}(z^p) dz= p \int_0^\infty z^{p-1} \omega_{|f|}(z)\;dz$ and $\{ |f|^p>z^p\}= \{|f|>z\}$. \qed \\




Ex: $f(x,y)=\dfrac{\sin xy}{(x^2+y^2)^{3/2}} \in L^1(E)$. $E= \{(x,y) \colon x,y>0\}$. 

Idea: estimate above by nonnegative function $g$ and prove $g \in L^1(E)$ by Tonelli. i.e. $|f| \geq g$. $|f| \geq \dfrac{1}{(x^2+y^2)^{3/2}}$ for sure but problems around origin but good at infinity. Have $|f| \leq \dfrac{xy}{(x^2+y^2)^{3/2}}$ good near 0. So use min of the two functions. Better: Use first when $x^2+y^2 \leq 1$: $g_1=\dfrac{1}{(x^2+y^2)^{3/2}} \chi_{\{x^2+y^2 \geq 1\}}$ and $\omega_{g_1}(\alpha)= \pi/(4\alpha^{2/3}) - \pi/4$ if $\alpha<1$ and 0 otherwise. Note that $1/r^3 \cdot \chi_{r \geq 1}$ and $\alpha<1$ $1/r^3\alpha$ and $r= 1/\alpha^{1/3}$ and $1/4$ if circle roots $1/\alpha^{1/3}$. 



%%%%%%%
% Friday

Silicing a subset of $\R^{n+m}$. Given a measurable set $E \subset \R^{n+m}$, consider $E_x= \{ y \in \R^m \colon (x,y) \in E \}$. 

% Blob with an x-slice and axes

\begin{itemize}
\item $E_x$ is measurable for almost all $x$.
\item If $|E|<\infty$, then $|E_x|<\infty$ for almost every $x$.
\item If $|E|=0$, then $|E_x|=0$ for almost every $x$. 
\end{itemize}

All these things follow easily from applying Tonelli to $\chi_E$ since almost every slice function is measurable. Slice fo $\chi_E$ is $\chi_{E_x}$ and $\int_{\R^{n+m}} \chi_E< \infty$ then $\int_{\R^n} |E_x| \; dx <\infty$.


\begin{ex} % Measurability in \R^{n+m}$ is important.
Assuming the Continuum Hypothesis, there exists $E \subset [0,1]^2$ such that every vertical slice $E_x$ is countable and every horizontal slice $E_y$ is co-countable, i.e. $[0,1] \sm$ countable set. Using TOnelli
	\[
	\int \left( \int \chi_E \; dy \right) \; dx = 0
	\]
since evert vertical slice is countable. The other way
	\[
	\int \left( \int \chi_E \; dx \right) \; dy = 1
	\]
This means $E$ is not measurable in $\R^2$ (otherwise these would be equal). Then let $E= \{ (x,y) \colon y \prec x \}$.
\end{ex}




Products:
If $E_1 \subset \R^n$, $E_2 \subset \R^m$, are measurable, then $E_1 \times E_2$ is measurable in $\R^{n+m}$ and $|E_1 \times E_2|= |E_1| \, |E_2|$, taking $0 \cdot \infty= 0$. Assume $E_1 \times E_2$ is measurable, then
	\[
	|E_1 \times E_2|= \int (\int \chi_{E_1 \times E_2}(x,y) \; dy) \; dx= \int |E_2| \; \chi_{E_1}(x) \; dx.
	\]
It remains to prove that $E_1 \times E_2$ is measurable. It suffices to prove if when $|E_1|<\infty$ and $|E_2|<\infty$. In the general case, $E_1= \cup_{k=1}^\infty E_{1,k}$ and $E_2= \cup_{k=1}^\infty E_{2,k}$, then $E_1 \times E_2= \cup_{j,k} E_{1,k} \times E_{2,j}$. 

There exists $G_1 \supset E_1$, $G_2 \supset E_2$ such that $|G_1 \sm E_1|<\ep$, $|G_2 \sm E_2|<\ep$. Now $G_1 \times G_2$ is open and $E_1 \times E_2 \subset G_1 \times G_2$. Now $(G_1 \times G_2) \sm (E_1 \times E_2) \subset [(G_1 \sm E_1) \times G_2] \cup [G_1 \times (G_2 \sm E_2)]$. Both $G_1 \sm E_1$ and $G_2 \sm E_2$ have measure at most $\ep$. Now $G_1, G_2$ by cumes $I_k, Q_j$ with summed measure at most $\ep$. Then $|I_k \times Q_j|= |I_k||Q_j|$ so $|(G_1 \sm E_1) \times G_2| < \sum_{k,j} |I_k| |Q_j| < \ep |G_2|$. Then just choose $\ep$ sufficiently small. 




Convolution: $(f*g)(x)= \int_{\R^n} f(x-t)g(t) \;dt$, where $f,g$ are defined on $\R^n$ are measurable.  Now $f*g=g*f$ by change of variable $s=x-t$ so $\int f(x-t)g(t) \; dt = \int f(s) g(x-s) \; ds$, where $s,t \in \R^n$. In the general case, $h= h^+ - h^-$. 

Claim $\int h(x-t) \; dt = \int h(t) \; dt$. Proof: if $h \geq 0$, then are measures of sets $\{(t,y) \colon 0 \leq y \leq h(x-t)\}$ and $\{ (t,y) \colon 0 \leq y \leq h(t) \}$. THe first is obtained from the other by reflection $t \mapsto -t$ and translation (add $x$). These do not change measure or measurability. 


\begin{ex}
Ex: Convolution: Argument of $\chi_{[a,b]} *\cdot \chi_{[c,d]}$ is only nonzero on $\{x \colon a \leq x-t \leq b, c \leq t \leq d\}$ so $\{x \colon x-b \leq t \leq x-a, c \leq t \leq d\}$. so on $\max(x-b,c) \leq t \leq \min(x-a,d)$. 
	\[
	\chi_{[a,b]} * \chi_{[c,d]}= \int_\R \chi_{[a,b]}(x-t) \chi_{[c,d]}(t) \; dt =  (\min(x-a,d) - \max(x-b,c))^+
	\]
So $\chi_{[-1,1]} * \chi_{[-1,1]}= (\min(x+1,1) - \max(x-1,-1))^+=$....grossly complicated (or just wrong). So better way $|([-b,-a] + x) \cap [c,d]|$ so $|([-1,1]+x) \cap [-1,1]|$.Then get 
% Graph, triangle with height 2 at x=0, at x-axis at -2,2, and 0 elsewhere, is the plot. 
\end{ex}


\begin{thm}
If $f,g \in L^1(\R^n)$, then $f*g \in L^1(\R^n)$ and $\int |f*g| \leq \int |f| \int |g|$.
\end{thm}

\pf We know $|f*g| \leq |f| * |g|$ by the triangle inequality. We can work with $|f|$ and $|g|$ so assume $f,g \geq 0$. $\int (\int f(x-t) g(t) \; dt) \; dx$ by tonelli this is $\int (\int f(x-t) g(t) \; dx ) \; dt = \int g(t) (\int f(x-t) \; dx \; )dt$ where the inner integral is just $\int f$ by translation. But then this final integral is $\int f \cdot \int g$. 

But why $f(x-t)g(t)$ measurable on $\R^{2n}$? (need this for tonelli). Observe $f(x)$ and $g(t)$ are measurable on $\R^{2n}$ since the preimages of $>$ something is just $\{(x,t) \colon f>a\}= \{x \colon f(x)>a\} \times \R$ since it does not depend on $t$. Then this is measurable. Use the linear transformation $\phi: (x,t) \to (x-t,x+t)$ intvertible inverse linear, hence lipschitz. Use then HW 3.5: composition $f \circ \phi$ is measurable, $\phi^{-1}$ Lipschitz. Hence $f(x-t)$ is measurable on $\R^{2n}$. 





%%%%%%%%%%%%

If $f,g$ are probability density functions of $X$ and $Y$, then $f*g$ is the PDF of $X+Y$, assuming $X,Y$ are independent. 

















% Ch 9. 4, p157

% ------------------------
%  Marcinkiewicz
% ------------------------
\subsection{Marcinkiewicz Integrals}

In Analysis, one often has $f*g$, where one of the functions is not in $L^1$. These are so called ``singular integrals.'' The simplest example is the Hilbert transform
	\[
	\int_\R \dfrac{f(x-t)}{t} \; dt
	\]

% Look this up and give better description.

Let $K$ be a compact set in $\R^n$, and $\delta(y):= \dist(y,k)= \min\{|x-y| \colon x \in K\}$. 

% Picture?


\begin{thm}
If $B \supset K$ is a bounded open set, then for all $\lambda > 0$. 
	\[
	\int \dfrac{\delta(y)^\lambda}{|x-y|^{n+\lambda}} \; dy < \infty
	\]
for a.e. $x \in K$.
\end{thm}

\pf It suffices to show $\int_K \mu_\lambda(x) \; dx < \infty$.
	\[
	\int_K \mu_\lambda(x) \; dx = \int_K \int_{B \sm K} \dfrac{\delta(y)^\lambda}{|x-y|^{n+\lambda}} \; dy \; dx = \int_{B \sm K} \int_K \dfrac{\delta(y)^\lambda}{|x-y|^{n+\lambda}} \; dx \; dy.
	\]
But now observe $|x-y| \geq \delta(y)$. But then 
	\[
	\int_{B \sm K} \int_K \dfrac{\delta(y)^\lambda}{|x-y|^{n+\lambda}} \; dx \; dy \leq \int_{B \sm K} \left( \int_{|x-y| \geq \delta(y)} \dfrac{\delta(y)^\lambda}{|x-y|^{n+\lambda}} \right) \; dy \leq \int_{B \sm K} C \; dx = C |B \sm K|< \infty.
	\]
Using $\int_{|x-y| \geq \delta} \dfrac{1}{|x-y|^{n+\lambda}} \; dx \leq \dfrac{C}{\delta^\lambda}$. To prove this:

Recall $\int |f|^p= p \int_0^\infty \alpha^{p-1} |\{|f|>\alpha\}| \; d \alpha= *$. Use $P= n + \lambda$, $f=\dfrac{1}{|x-y|}$.
	\[
	| \{ |x-y| \geq \delta \colon \dfrac{1}{|x-y|}>\alpha\}| = |B(y,1/\alpha) \sm B(y,\delta)|= C((1/\alpha)^n - \delta^n)^+
	\]
which is $=0$ if $\alpha>1/\delta$. For *, note
	\[
	*= (n+\lambda)^{1/\delta} \alpha^{n+\lambda-1} C (1/\alpha^n - \delta^n) \; d\alpha \leq C' \int_0^{1/\delta^n} \alpha^{\lambda-1} \; d \alpha = C'' \dfrac{1}{\delta^\lambda}
	\]





$B$ is needed as the integral is $\sim \dfrac{1}{|y|^n}$ as $y \to \infty$. We could have integrate over $B \sm K$ since $\delta \equiv 0$ on $K$. Also, the $\delta$ is important for convergence. The almost always is also unavoidable: take $K=\{0\}$ in $\R$. Then $\delta(y)= |y|$. Then $\int_B \dfrac{|y|^\lambda}{|x-y|^{n+\lambda}} \; dy$ so at $x=0$, we get $\int_B \dfrac{1}{|y|^n}= \infty$. Generally, $\mu_\lambda(x)= \infty$ if $x$ is an isolated point of $K$. If $x$ is an interior point fo $K$, $\mu_\lambda(x)< \infty$ as $\delta=0$ in a neighborhood of $x$.




Question: in last part of proof, can we let $\lambda=0$? Specifically, is it true that $\int_{B \sm K} \dfrac{1}{|x-y|^n} \; dy< \infty$ for a.e. $x \in K$? Counterexample? $K$ is fat Cantor set?




















