% !TEX root = ../../mat701_notes.tex
\newpage
\section{Abstract Integration}
\subsection{Measure Spaces}

So far, we only have had Lebesgue measure on $\R^n$. But there are other measures: length of curves in $\R^2$, surface area on surfaces, probability on probability space $\Omega$ $P(\text{event})=$ measure of outcomes. 

% Give weird closed curve
% Give picture of torus
% Give bulls eye in circle. 

\begin{dfn}[Measure Space]
A measure space is a triple, $(X,\Sigma,\mu)$, where $X$ is a set, $\Sigma$ is a $\sigma$-algebra on subsets of $X$, and $\mu: \Sigma \to [0,\infty]$ is a countably additive function on disjoint sets. 
\end{dfn}


\begin{rem}
One often also requires $\mu(\emptyset)=0$. However, this does almost follow from our definition since $\mu \geq 0$ and if $\emptyset \neq A \subseteq X$, then 
	\[
	\mu(A)=\mu(A \cup \emptyset)= \mu(A) + \mu(\emptyset)
	\]
so that $\mu(\emptyset)=0$ unless $\mu(\emptyset)=\infty$.  
\end{rem}

\begin{ex} \hfill
\begin{enumerate}[(i)]
\item $(\R^n,\mathcal{L},\mu)$, $\mathcal{L}$ is the the set of Lebesgue measurable set. $\mu(E)=|E|$.

\item $(X,2^X,\mu)$, $X$ is any set, $\mu(E)$ is the cardinality of the set. 
\item Like (ii) but
	\[
	\mu(E)=
	\begin{cases}
	0, & E \text{ countable} \\
	\infty, & E \text{ uncountable}
	\end{cases}
	\]

\item $(\R^n,L,\mu_w)$, where
	\[
	\mu_w(E)= \int_E w
	\]
where $w \geq 0$ is measurable. One may say that $w$ is the `weight' or `density.'  % Gaussian curve
\end{enumerate}
\end{ex}


\begin{nex}
like (ii) above but
	\[
	\mu(E)=
	\begin{cases}
	0, & |E|<\infty \\
	\infty, & |E| \text{ infinite}
	\end{cases}
	\]
this is \emph{not} a measure space since it is not countably additive. 
\end{nex}


\begin{dfn}[Additive Set Function (ASF)]
We say a function $\phi: \Sigma \to \R$, where $\Sigma$ is a $\sigma$-algebra, is countably additive. 
\end{dfn}


The difference between $\phi$ and a measure is that it can be negative and it cannot be infinite. This is also called a \emph{finite signed measure}. 


\begin{ex}
$\phi(E)= \int_E f$, where $f \in L^1$. 
\end{ex}

An advantage of ASF can pass to complement: $\phi(E^C)= \phi(X) - \phi(E)$ and get results for $\cap$ from $\cup$.


\begin{thm}
ASF is continuous on monotone sequences of sets, i.e. $E_1 \subset E_2 \subset \cdots$ then $\phi(\cup E_k)= \lim \phi(E_k)$ or $E_1 \supset E_2 \supset \cdots$ then $\phi(\cap E_k)= \lim \phi(E_k)$
\end{thm}

\pf The second follows from the first by taking complements. Now
	\[
	\bigcup_{k=1}^\infty E_k= E_1 \cup (E_2 \sm E_1) \cup (E_3 \sm E_2 \sm E_1) \cup \cdots
	\]
disjoint union so
	\[
	\phi(\cup E_k) = \phi(E_1) + (\phi(E_2) - \phi(E_1)) + \cdots
	\]
which telescopes so we have $\lim_{k \to \infty} \phi(E_k)$. \qed \\


For measures $\mu$. the first works but in the second statement above, we require $\mu(E_1)< \infty$.


\begin{thm}[`Fatou's Lemma']
if $\phi \geq 0$ is an ASF (i.e. finite measure), then $\phi(\liminf E_k) \leq \liminf \phi(E_k)$. Hence by considering $E^C_k$, $\phi(\limsup E_k) \geq \limsup \phi(E_k)$. 
\end{thm}

\pf Let $M_k= \bigcap_{j \geq k} E_j$, then $\liminf E_k= \bigcup_{k=1}^\infty H_k$.
	\[
	\phi \left( \bigcup H_k \right) = \lim_{k \to \infty} \phi(H_k)= \liminf_{k \to \infty} \phi(E_k),
	\]
using $H_k \subset E_k$, so $\phi(E_k) - \phi(H_k)= \phi(E_k \sm H_k) \geq 0$. 




Jordan Decomposition for ASF 

Every ASF is the difference of two nonnegative ASFs. 

Application: Write ASF $\phi = \phi_1 - \phi_2$, both nonnegative. Conclude if $E= \limsup E_k= \liminf E_k$, then $\phi(E)= \lim \phi(E_k)$. 


\begin{dfn}[Upper Variation]
Given an ASF $\phi$, upper variation $\overline{V}(E)= \sup_{A \subset E} \phi(A)$. 
\end{dfn}

\begin{dfn}[Lower Variation]
Given an ASF $\phi$, lower variation $\underline{V}(E)= - \inf_{A \subset E} \phi(A)$. 
\end{dfn}

\begin{dfn}[Total Variation]
Given an ASF $\phi$, total variation: $V(E)= \overline{V}(E) + \underline{V}(E)$.
\end{dfn}

How does this relate to Jordan decomposition? Idea of proof is to show $\phi(E)= \overline{V})(E) - \underline{V}(E)$. 

where (needs to be proved) $\overline{V}$, $\underline{V}$ are ASFs. Why this was
	\[
	\int_{A \subset E, A \in \Sigma} \phi(A) = \inf_{B \subset E, B \in \Sigma} (\phi(E) - \phi(B)) = \phi(E) - \sup_{B \subset E, B \in \Sigma} \phi(B) = \phi(E) - \overline{V}(E)
	\]
Take $A= E \sm B$. Hence, $\phi(E)= \overline{V}(E) - \underline{V}(E)$. 


Clear total variation is ASF if the upper/lower are. So sketch proof that they are:

1) Countable subadditivity. 
$A \subset \cup E_k$, write as $A= (A \cap E_1) \cup (A \cap E_2 \sm E_1) \cup \cdots$. Get $A= \cup A_k$, $A_k \subset E_k$. $\phi(A_k) \leq \overline{V}(E_k)$ hence $\phi(A) \leq \sum \overline{V} (E_k)$. The upper follows the same.

2) $V< \infty$, which is done by contradiction. 
$\overline{V}(E) \geq \sum \overline{V}(E_k)$ just pick $A_k \subset E_k$, $\phi(A_k) \geq \overline{V}(E_k) - \ep/2^k$, use countable subadditivity of $\phi$. 




%%%%%%%%%%%%%%%%%%%%%%


\subsection{Measurable Functions and Integrals}

Repeat all previous work for abstract measure spaces. Recall $(X, \Sigma, \mu)$ is a measure space. 

\begin{dfn}
A function $f: X \to \overline{\R}$ is measurable if $\{f>a\} \in \Sigma$ for all $a \in \R$. 
\end{dfn}

Fur functions $f: X \to \R$, equivalent to $f^{-1}(\text{Borel}) \in \Sigma$. 

\begin{ex}
Let $X= \N$, $\Sigma = 2^X$, then all functors are measurable.
\end{ex}

Basic properties: $cf$, $f+g$, $\sup_k f_k$, $\inf_k f_k$, $\lim_k f_k$ are measurable. 

$\sum_{\text{finite}} a_k \chi_{E_k}$, disjoint, $a_k$ unique is measurable if and only if $E_k \in \Sigma$. 

Every nonnegative measurable function $f$ is the limit of $f_k$ (simple), $f_k \nearrow f$, $f_k = \min(k, 2^{-k} \text{ floor}(2^k f)$, second piece round down to $j/2^k$: $\{f_k = j/2^k\} = \{ j/2^k \leq f< (j+1)/2^k\}$ or $\emptyset$. 

What does not \emph{not} generalize.

``continuous'' then measurable. Continuity is not even defined on $X$.

Lusin's Theorem does not apply.

If $f$ is measurable and $f=g$ a.e. wrt $\mu$ then $g$ not necessarily measurable. (because subsets of measure zero no longer necessarily measurable). example: $(\R, \mathcal{B}, |\cdot|)$, Borel with Lebesgue measure. in Cantor set $C$, there is a non-Borel subset $A$. $\chi_A \equiv 0$ a.e. but $\chi_A$ is not measurable. 


$X= \bigcup_{k=1}^\infty E_k$, $E_k$ has finite measure. Example: $\R$ with counting measure. ie $\infty$ on infinite sets and finite otherwise. 


\begin{dfn}
If $f \geq 0$ is measurable. $\int_E f \; d\mu = \sup\{ \sum_j a_j \mu(E_j) \}$ over simple functions $\sum a_j \chi_{E_j} \leq f \}$. 
\end{dfn}

Can choose $a_j = \inf_{E_j} f$, so $\int f \; d\mu= \sup_p L(f,P)$, where $P$ is partition of $E$ into $E_1, \ldots, E_m$, $L(f,P)= \sum_j \int_{E_j} f \mu(E_j)$. Choosing finer partition $P$, get larger $L(f,P)$. 


\begin{dfn}
If $f$ measurable and one of $\int_E f^+ \; d\mu$, $\int_E f^- \; d\mu$ is finite, define $\int_E f \; d\mu= \int_E f^+ \; d\mu - \int_E f^- \; d\mu$. if both finite, $f \in L^1(E,\mu)$. 
\end{dfn}


\begin{ex}
$(\N,2^\N,\text{counting})$, $\int f \;d\mu= \sum_{k=1}^\infty f(k)$, if abs convergent. $L(f,P)= \sum_{k=1}^N f(k)$ So integrals are series! Notice integral finite iff abs convergent. 
\end{ex}


\begin{ex}
$(\R, 2^\R, \delta_{x_0})$. $\delta_{x_0}$ the point mass at $x_0$, i.e. dirac delta 
	\[
	\delta_{x_0}= 
	\begin{cases}
	1, & x_0 \in A \\
	0, & x_0 \notin A
	\end{cases}
	\]
$\int_\R f \; d\delta_{x_0}= f(x_0)$ \pf Suppose $f$ nonnegative (otherwise, use $f^+$ and $-f^-$ and then combine later). Let $P=\{E_1,\ldots,E_m\}$ be a partition. One $E_j$ contains $x_0$, then $L(f,P)= \inf_{E_j} f \leq f(x_0)$. Hence, $\int f \leq f(x_0)$. Also, $P = \{ \{x_0\}, \R \sm \{x_0\} \}$ gives $L(f,P)= f(x_0)$. 
\end{ex}


We need another proof of MCT, since before used ``volume definition'' of $\int f$ and this does not generalize. 


\begin{lem}
if $E= A \cup B$, disjoint measurable, then $\int_E f= \int_A f + \int_B f$ for nonnegative measurable $f$. 
\end{lem}

\pf Any partition $P$ of $E$ can be refined, $(E_j$ replaced by $E_j \cap A$, $E_j \cap B$) so we get $P= P_A \cup P_B$. Then $L(f,P)= L(f_1,P_A) + L(f,P_B)$. Hence, $\int f \leq \int_A f + \int_B f$. But $P_A,P_B$ approx $\int_A f, \int_B f$ part them into $P$, conclude $\int_E f \geq \int_A f + \int_B f$. \qed \\


%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Continuing 10.2

\begin{ex}
$X= \R^2$, $\Sigma=$ Borel sets, $\mu(E)= | \{x \in \R \colon (x,0) \in E\}|$. $\int f \; d\mu= \int_\R f(x,0) \; dx$. Important example of integrating along a curve. 
\end{ex}


Our goal is now to prove the MCT (if $f_k \nearrow f$, $f_k \geq 0$ measurable, then $\int f_k \to \int f$). We will need some preparation. 

Subadditivity over $f$: $\int (f+g) = \int f + \int g$, where $f,g \geq 0$ measurable. How do we prove this? For simple, easy: $f= \sum a_k \chi_{E_k}$, $g= \sum b_l \chi_{F_l}$, rewrite as $\sum_{k,l} a_k \chi_{E_k \cap F_l}$ and $\sum_{k,l} b_l \chi_{E_k \cap F_l}$. The first is integral $\sum a_k \mu(E_k \cap F_l)$ and second is integral $\sum b_l \mu(E_k \cap F_l)$. But then for the sum: $\sum (a_k+b_l) \chi_{E_k \cap F_l}$ converges to integral $\sum (a_k+b_l) \mu(E_k \cap F_l)$.


\begin{lem}
Suppose $f_k, g \geq 0$ are simple, $f_1 \leq f_2 \leq \cdots \lim f_k \geq g$. Then $\lim \int f_k \geq \int g$.
\end{lem}

\pf Partition set $E$ to make $g$ constant (on each piece) $g \equiv c$. Given $c' < c$. Let $A_k = \{f_k >c \}$. Then $A_1 \subset A_2 \subset \cdots$ and $\bigcup A_k = E$, so $\mu(A_k) \to \mu(E)$. But $\int_E f_k \geq \int_{A_k} f_k \geq c' \mu(A_k)$. So $\lim \int f_k \geq c' \mu(E)$. So $\lim \int f_k \geq c \mu(E)= \int g$. \qed \\


\begin{lem}
If $f_k \nearrow f$, $f_k$ simple nonnegative, then $\int f_k \to \int f$. 
\end{lem}

\pf For all simple $g \leq f$, we have $\lim \int f_k \geq \int g$ by the previous lemma. Take $\sup$ over $g$, get $\lim_l \int f_k \geq \int f$. But reverse is immediate since $f_k \nearrow f$ so $f_k \leq f$. \qed \\


Proof of $\int (f+g) = \int f + \int g$.

Take simple $f_k \nearrow f$, $g_k \nearrow g$. Pass to limit in $\int (f_k + g_k) = \int f_k + \int g_k$. \qed \\

Then follows for general function by writing $f^+$ and $f^-$. 


Monotone Convergence Theorem (Two forms):

Sequence Form: if $f_k$ are nonnegative, measurable, and $f_k \nearrow f$, then $\int f_k \to \int f$.

Series Form: If $g_k \geq 0$ measurable, then $\in \sum g_k = \sum \int g_k$.


These are equivalent. [The duality of sequences and series in a sense.]

\pf (Series Form since easier to prove) $\int g \geq \sum_{k=1}^\infty \int g_k$ by considering $g \geq \sum_{k=1}^N g_k$. $\int g \geq \sum_{k=1}^N \int g_k$. Need $\leq$. Introduce simple $g_{k,j} \nearrow g_k$ as $j \to \infty$.
	\[
	\begin{split}
	g_{11} \quad g_{12} \quad g_{13} \quad \cdots \to g_1 \\
	g_{21} \quad g_{22} \quad g_{23} \quad \cdots \to g_2 \\
	g_{31} \quad g_{32} \quad g_{33} \quad \cdots \to g_3 \\
	\end{split}
	\]
% Give box around upper left corner, then top to on second column, top three on third columnm, etc.

We use an ``upper triangular sum''. $s_j = \sum_{k=1}^j g_{k,j}$ which converges monotonically to $g$, i.e. $s_j \nearrow g$. By second lemma above, $\int s_j \to \int g$. Also, $\sum_{k=1}^j g_k \geq s_j$, hence $\sum_1^\infty \int g_l \geq \int g$. \qed \\



MCT implies Fatou which implies DCT.

H\"older, Minkowski: $L^p(E,\mu)$ is a Banach space. 






% ------------------------
%  Absolute Continuity
% ------------------------
\subsection{Absolute Continuity \& Singular Sets}
Now we move on to 10.3 


Given a measure space $(X, \Sigma, \mu)$ and an ASF $\phi$, we say

\begin{dfn}[Absolutely Continuous]
$\phi$ is absolutely continuous with respect to $\mu$ if $\mu(A)=0$, then $\phi(A)=0$. 
\end{dfn}


\begin{dfn}[Singular]
$\phi$ is singular with respect to $\mu$ if there exists $Z$ such that $\mu(Z)=0$ and $\phi(A)= 0$ for all $A \subset Z^C$. 
\end{dfn}


Equivanetly, for total variation $V$

AC: $\mu(A)=0$ then $V(A)=0$

S: $\mu(Z)=0$ and $V(Z^C)=0$.

Main results: every ASF is the sum of AC and S. AC ASF has density $\phi(A)= \int_A \omega$.

Examples:

a) $\mu=$ Lebesgue measure on $\R$ and $\phi(A)= \int_A f$, $f \in L^1(\R)$. THis is AC

b) same $\mu$, $\phi(A)$ is 1 if $x_0 \in A$ and $0$ if $x_0 \notin A= \delta_{x_0}$. 

Singular: let $Z= \{x_0\}$ then $\mu(Z)=0$, $V(Z^C)=0$. 


On $[0,1]$, fix $w \geq 0$ integrable. 
fix $w \geq 0$, $w \in L$, $\mu_w(A) = \int_A w$.

Is lebesuge measure on $[0,1]$ AC wrt $\mu_w$? No if $|\{w=0\}|>0$ 

Yes if $w>0$ a.e.

% Plot, axes, function w, along x-axis then linearly increasing for a bit. thick black lines, tick mark along x-axis where the linear function stops. 



%%%%%%%%%%%%%%%%%%%%%%%%%%%


Goal: Every ASF is AC $+$ S: AC has `density' $f$, $\phi(A)= \int_A f$. 



Remark: for any $E \subset \Sigma$, $(E, \Sigma'= \{F \in \Sigma \colon F \subset E\}, \mu \big|_{\Sigma''})$ is a measure space (subspace of $X$). 


\begin{thm}
(From $0-0$ to $\ep-\delta$ in AC) $\phi$ is AC if and only if for all $\ep>0$, there exists $\delta>0$ such that for all $A$ $\mu(A)<\delta$ then $|\phi(A)|<\ep$.
\end{thm}

\pf Recall $\phi= \ov{V} - \underline{V}$, so $|\phi(A)| \leq V(A)$. Suppose there exists $\ep>0$ for all $k$ there exists $A_k$ with $\mu(A_k)< 2^{-k}$ but $V(A_k) \geq \ep$. Let $A= \limsup A_k$. Then $\mu(A)=0$ (because $\sum \mu(A_k)< \infty$) but $V(A) \geq \limsup V(A_k) \geq \ep$. (flipped Fatou). This contradicts AC. \qed \\


Ex: For infinite measures $\mu, \nu$, this is not true. For example, $\mu$ is the Lebesgue measure on $\R$ and $\nu(A) = \int_A x^2 \; dx$. $\nu$ is AC wrt $\mu$. If $\mu(A)=0$, then $\nu(A)=0$ but for all $\delta>0$, $\int_{[n,n+\delta]} x^2 \; dx \to \infty$ as $n \to \infty$. 


Hahn decomposition: if $\phi$ is ASF, then there exists ``positive set'' $P$ such that $\phi \geq 0$ on subsets of $P$ and $\phi \leq 0$ on subsets of $P^C$.

\pf $P$ should maximize $\phi$. Recall $\ov{V}(X)= \sup \{ \phi(A) \colon A \in \Sigma\}$ for all $k$. Choose $A_k$ with $\phi(A_k) > \ov{V}(X) - 2^{-k}$. Let $P= \liminf A_k$. Note $\underline{V}(A_k)= \ov{V}(A_k) - \phi(A_k)$. Fatou's lemma gives $\underline{V}(A) \leq \liminf \underline{V}(A_k)= 0$. Also, $P^C= \limsup A_K^C$. and $\ov{V}(A_K^C)= \ov{V}(X) - \ov{V}(A_k)< 1/2^k$. Hence, $\ov{V}(P^C)=0$ [$\sum \ov{V}(A_k^C)<\infty$] \qed \\


Ex: if $\phi(A)= \int_A f$, then $P= \{ f \geq 0 \}$ or $P=\{ f>0 \}$. 

Compare to Jordan decomposition: $\phi= \ov{V} - \underline{V}$. Hahn makes this more precise. $\ov{V}(A)= \phi(A \cap P)$, $-\underline{V}(A)= \phi(A \sm P)$ so $\underbrace{\phi(A \cap P)}_{\ov{V}} - \underbrace{(- \phi(A \sm P))}_{\underline{V}}$.


\begin{lem}[Level Decomposition of ASF]
Suppose $\mu(X)< \infty$, $\phi \geq 0$ an ASF, then for all $a>0$, there exists partition $X= Z \cup \bigcup_{k=1}^\infty E_k$, $\mu(Z)=0$
	\[
	a(k-1) \mu(A) \leq \phi(A) \leq a k \mu(A)
	\]
for all $A \subset E_k$
	\[
	a(k-1) \leq \dfrac{\phi}{\mu} \leq ak
	\]
on $E_k$
\end{lem}


Ex: if $\phi(A) = \int_A \dfrac{1}{\sqrt{x}} \; dx$ on $X=[0,1]$ $a=1$ $Z= \emptyset$, $E_1= \emptyset$

% Give plot focused on Q1 part, E_2, E_3, E_4 moving from right to left. 

\pf WLOF, $a=1$ (else consider $\phi/a$). Apply Hahn to $\phi - k \mu$ ($k \in \N$) get a set $P_k$, $\phi \geq k \mu$ on $P_k$ and $\phi \leq k \mu$ on $P_k^C$. Let $Q_k= \bigcup_{m \geq k} P_m$ note $\phi \geq k \mu$ on $Q_k$, $\phi \leq k \mu$ on $Q_k$. Let $Z= \bigcap_k Q_k$ on this set $\phi \geq k \mu$ so $\mu(Z)=0$. Let $E_1= Q_1^C$ then $\phi \leq \mu$ on $E_1$, for $k>1$, let $E_k = Q_{k-1} \sm Q_k$, then $(k-1) \mu \leq \phi k \mu$ as required. \qed \\




% Check about finiteness for theorem? Needed? How to state?

\begin{thm}[Lebesgue Decomposition]
Let $\phi$ be an ASF and $\mu$ a $\sigma$-finite measure. Then there exists a unique additive set function $\alpha,\sigma$ such that $\phi = \alpha + \sigma$, $\alpha$ is AC wrt $\mu$, $\sigma$ is S wrt $\mu$. Moreover, there exists $f \in L^1(X,\mu)$ such that $\alpha(A)= \int_A f$ and there exists $Z$ such that $\mu(Z)=0$ and $\sigma(A)= \phi(A \cap Z)$.
\end{thm}

\pf Uniqueness: $\alpha+\sigma= \alpha' + \sigma'$, then $\alpha-\alpha'= \sigma'-\sigma$ so then both sides AC and S by equality, hence 0. 



Existence of $Z$: comes from definition of singular. ASF $\mu(Z)=0$ and $\phi=0$ on $Z^C$. So $\sigma(A)= \sigma(A \cap Z)$. But $\alpha(A \cap Z)=0$ because $\alpha$ is AC. Hence, $\sigma(A)= \phi(A \cap Z)$. 

Hard parts existence of decomposition and existence of $f$. 




%%%%%%%%%%%%%%

Why do we care? Duality of functions spaces such as $L^p$. Dual space of $L^p$ is $\{ \psi : L^p \to \C \colon \psi \text{ linear cont. }\}$. $\psi$ induces ASF $\phi(E)= \psi(\chi_E)$, $\phi(E \cup F)= \psi(\chi_E + \chi_F)= \phi(E) + \phi(F)$ if disjoint. 


Why need to be $\sigma$-finite?  ($X= \bigcup_{k \in \N} E_k$, $\mu(E_k)<\infty$)

\begin{ex}
Let $\mu$ be counting measure on $[0,1]$ (gives number of elements of set). $X=[0,1]$. $\phi(A)= |A|$ is an ASF. If $\phi= \alpha + \sigma$, then $\sigma(A)=0$ for all $A$ (because the only $Z$ with $\mu(Z)=0$ is $\emptyset$). Hence, $\phi(A)=  \int_A f \; d\mu$ for some $f \in L^1(\mu)$, which means $\sum_{x \in [0,1]} |f(x)|<\infty$. But then $f \equiv 0$ except on a countable set $C$. But then get $\phi([0,1] \sm C)= 0$, contradicting $\phi(A)=|A|$.
\end{ex}


\pf (LDT) Can reduce to $\phi \geq 0$. For general case $\phi$, $\phi = \ov{V} - \underline{V}$ so if $\ov{V} = \alpha_1 + \theta_1$, $\underline{V}= \alpha_2 + \theta_2$, then $\phi= (\alpha_1-\alpha_2) + (\theta_1-\theta_2)$.

2) Reduce to $\mu(X)<\infty$. Else $X= \cup_{\text{disjoint}} E_k$, If $\phi_k(A)= \phi(A)$ for $A \subset E_k$ has LD, $\alpha_k+\theta_k$, then let $\alpha(A)= \sum \alpha_k(A \cap E_k)$, $\sigma(A)= \sum \sigma_k(A \cap E_k)$. 

3) Main case: $\phi \geq 0$, $\mu(X)< \infty$. Use level decomposition with $a= 2^{-m}$. Obtain $X= Z^m \cup \bigcup_{k=1}^\infty E_k^m$, $\mu(Z^m)=0$ and $\phi(A)/\mu(A)$ is in $[(k-1)2^{-m}, k2^{-m}]$ for all $A \subset E_k^m$. Let $Z = \bigcup_{m=1}^\infty Z^m$, this will be our $Z$. Need $\phi(A \sm Z)= \int_A f \; d\mu$. Let $f_m = \sum_k (k-1) 2^{-m} \chi_{E_k^m}$. How close is $\int_A f_m$ to $\phi(A \sm Z)$?

$\phi$ is AC wrt $\mu$ on each $E_k^m$, hence $\phi(A \cap E_k^m)= \phi((A \sm Z) \cap E_k^m)$, null subsets of $E^m$ are negligible. So $\phi(A) = \phi(A \cap Z) + \phi(A \sm Z) = \phi(A \cap Z) + \sum_k \phi(A \cap E_k^m)$. 


$\phi(A \cap E_k^m) \geq (k-1) 2^{-m} \mu(A \cap E_k^m)= \int_{A \cap E_k} f_m$. $\phi(A \cap E_k^m) \leq k 2^{-m} \mu(A \cap E_k^m)= \int_{A \cap E_k^m} f_m + 2^{-m} \mu(A \cap E_k^m)$. Sum over $k$
        \[
        0 \leq \phi(A \sm Z) - \int_a f_m \; d\mu \leq 2^{-m} \mu(X)
        \]
Remains to show $f_m$ converge uniformly as $m \to \infty$ (then $\phi(A \sm Z)= \int_A f \; d\mu$ with $f= \lim f_m$). 


Observe $E_k^m$ is covered by $E_j^{m+1}$ for $j= 2k-2$, $2k-1$, $2k$, $2k+1$ because
	\[
	\left[ \dfrac{k-1}{2^m}, \dfrac{k}{2^m} \right] \bigcap \left[ \dfrac{j-1}{2^{m+1}}, \dfrac{j}{2^{m+1}} \right]= \emptyset
	\]
unless $j$ is one of the values above. So if $f_m = (k-1) 2^{-m}$ then $f_{m+1}= j 2^{-m-1}$, where $2k-2 \leq j \leq 2k+1$. hence $|f_m- f_{m+1}| \leq 2^{-m}$. We get uniformly convergent of $\{f_m\}$. \qed \\



Radon-Nikodym's Theorem:

Every AC ASF $f$ wrt to $\sigma$-finite measure is of the form $\phi(A)= \int_A f \; d\mu$ for some $f \in L^1(\mu)$. 

% Look up better statement.


Lebesgue decomposition for measures: Let $\mu, \nu$ be $\sigma$-finite measures on $(X,\Sigma)$. Then there exists measurable $f \geq 0$ and a measure $\sigma$ such that $\sigma$ is singular with respect to $\mu$. $\nu(A)= \int_A f \; d\mu + \sigma(A)$.

\pf Let $X= \bigcup E_j$, disjoint. $\nu(E_j)< \infty$. $\nu \big|_{E_j}$ is an ASF, so $\nu(A)= \int f_j \; d\mu + \sigma_j(A)$ for $A \subset E_j$, $f_j \in L^1(E_j, \mu)$. Note $f_j \geq 0$.

Let $\sigma(A) = \sum_j \sigma(A \cap E_j)$ and $f= \sum f_j \chi_{E_j}$ (gluing). Hence $\nu(A)= \sum_j \nu(A \cap E_j)= \sum_j (\int_{A \cap E_j} f_j \; d\mu+ \sigma_j(A \cap E_j))= \int_A f \; d\mu + \sigma(A)$. \qed \\



%%%%%%%%%%

% 10.4

\subsection{The Dual Space of $L^p$}

% Remark on the Cantor measure from HW. 
%K=[0,1], \mu_0 = Lebesuge |_{[0,1]} restricted measure to A is \mu|_A, \mu|_A(E)= \mu(A \cap E), K_1, \mu_1= 3/2 \cdot Lebesgue|_{K_1}, \mu_1(E)= 3/2 |E \cap K_1|
%K_n 2^n intervals of length 3^{-n}. \mu_n(3/2)^n Lebesgue|_{K_n} each of 2^n intervals gets 2^{-n}, limit? get \mu such that \mu(C)=1 and \mu(\R\smC)=0. Note \mu_n(C)=0 for all N but \mu(C)=1, so a different kind of convergence is required. 
% Ways to construct \mu:
%1. weak convergence of measures \mu_n \ma{weak^*} \mu (702)
%2. product measures: \{0,1\}^\infty= all binary sequences. \{0,1\} has \nu: \nu(\{0\})=1/2, \nu(\{1\})=1/2, \prod_{k=1}^\infty \nu on \prod_{k=1}^\infty \{0,1\} is the Cantor measure
%3. Use outer measure: `hausdorff measure of dimension log 2/log 3, need ch 11
%4. HW \mu= pushforward under certain f of lebesgue.


\begin{dfn}
Let $B$ a normed vector space over $\R$. Define
	\[
	\begin{split}
	B^*&= \{ \text{ linear } \psi: B \to \R \colon \exists c \; |\psi(v)| \leq c \|v\| \forall v \in B \} \\
	&= \{ \text{ bounded linear functionals} \}
	\end{split}
	\]
is the dual space of $L^p$.
\end{dfn}


\begin{thm}
Let $(X,\Sigma,\mu)$ be a $\sigma$-finite measure space, $1 \leq p < \infty$. Then there is a linear bijection between $L^p(X,\mu)^*$ and $L^{p'}(X,\mu)$ which preserves the norm. 
\end{thm}

Specifically, the bijection sends $g \in L^p$ to $\psi_g \in (L^p)^*$, $\psi_g(f)= \int_X fg \; d\mu$. 


\begin{ex}
$(X,\Sigma,\mu) = (\N, 2^\N, \text{counting})$, so $L^p(X,\mu)$ is $\ell^p$: every $\psi: \ell^p \to \R$ is $\psi(x)= \sum_{k=1}^\infty x_k y_k$ for some $y \in \ell^{p'}$. Note so for $\ell^\infty$. One can define $\psi(x)= \lim_{k \to \infty} x_k$ when exists, and extend to a bounded linear functional on $\ell^\infty$. Suppose we had $y \in \ell^1$ such that $\sum_1^\infty x_ky_k = \lim_{k \to \infty} x_k$ for all convergent sequences $x$. Problem is limit does not depend on any particular $x_k$. So can choose one and add one to it, etc. so we must have $y_k=0$. But this is true for all of them so that we must have $y=0$ but then if $x=1$ for all $k$ so sum is not equal to the limit, which is 1, a contradiction. 
\end{ex}


\pf If $g \in L^{p'}$, then $|\int fg| \leq \|f\|_p \|g\|_{p'}$, so $\psi_g$ is bounded: $|\psi_g(f)| \leq \|g\|_{p'} \|f\|_p$. So $\|\psi_g\| \leq \|g\|_{L^{p'}}$. This gives a linear map $L^{p'} \to (L^p)^*$. This map is injective: if $\int fg=0$ for all $f \in L^p$, then $g \equiv 0$ a.e., LTR. Surjectivity: $\sigma$-finite case can be reduced to $\mu(X)<\infty$ by $X= \cup_1^\infty E_j$, $E_1 \subset E_2 \subset E_3 \subset \cdots$. 

% Sometimes useful to write as union finite sets, someitmes better to write as chian of increasing sets in measure, here this is the case.

We get $g_j$ on $E_j$ from the finite case ($\mu(E_j)<\infty$). Glue them into one $g$ which gives $\psi(f)= \int_X fg$. Finite case: $\mu(X)< \infty$. Define an ASF $\phi(A)= \psi(\chi_A)$. It is AC wrt $\mu: \mu(A)=0$ measure $\chi_A=0$ a.e., so $\psi(\chi_A)=0$. By Radon-Nikodym, there exists $g \in L^1(X,\mu)$ such that $\phi(A)= \int_A g \; d\mu$ for all $A \subset X$. This means $\psi(\chi_A)= \int g \chi_A \;d\mu$. $\psi(f)= \int gf \; d\mu$ for all simple $f$. Once we know that $g \in L^p$, we can conclude the proof as follows: $\psi$ and $\psi g$ are continuous on $L^p$ and $\psi= \psi g$ on a dense subset (simple functions), hence $\psi = \psi g$. 

case $p=1$, $p'=\infty$. Case $1<p<\infty$, $1<p'<\infty$ (textbook). for the first then, $\phi(A) = \psi(\chi_A) \leq \|\psi\| \cdot \|\chi_A\|_1 = \|\psi\| \mu(A)$. Hence, $\phi(A)/\mu(A) \leq \|\psi\|$. In the proof of Lebesgue decompostion, we will not have density $>\|\psi\|$ because of the inequality just shown. Thus $\|g\|_{p'} \leq \|\psi\|$. \qed \\





































